---
title: "Homework 01"
---

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
```


```{r}
#Method extracted from the repository
tuesdata <- tidytuesdayR::tt_load('2026-01-27') 
```

```{r}
#Method extracted from the repository
companies <- tuesdata$companies
legal_nature <- tuesdata$legal_nature
qualifications <- tuesdata$qualifications
size <- tuesdata$size
```


## TidyTuesday Section

Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.



*Research question:*
How does the distribution of capital stock among Brazil’s top 275 companies vary by company size and legal nature?

```{r}
# Cleaning the data 
top_companies <- companies %>%
  arrange(desc(capital_stock)) %>%  
  slice_head(n = 275)
```


```{r}
# Company size 
company_size <- top_companies %>%
  group_by(company_size) %>%
  summarise(capital = sum(capital_stock, na.rm = TRUE)) %>%
  mutate(total_share = capital / sum(capital) * 100) %>%
  arrange(desc(total_share))
```


```{r}
# Legal nature 
legal_nature <- top_companies %>%
  group_by(legal_nature) %>%
  summarise(capital = sum(capital_stock, na.rm = TRUE)) %>%
  mutate(total_share = capital / sum(capital) * 100) %>%
  arrange(desc(total_share))
```



*Visualizations*


```{r}
ggplot(company_size, aes(x = total_share, y = reorder(company_size, total_share))) +
  geom_col(fill = "blue") +
  labs(title = "Capital Share by Company Size", x = "Share (%)", y = "Company Size") +
  theme_minimal()
```

```{r}
ggplot(legal_nature, aes(x = total_share, y = reorder(legal_nature, total_share))) +
  geom_col(fill = "green") +
  labs(
    title = "Capital Share by Legal Nature",
    x = "Share (%)",
    y = "Legal Nature"
  ) +
  theme_minimal()
```



