---
title: "Homework 06"
format: 
  html: 
    embed-resources: true
---

## TidyTuesday Section (optional)

::: {.callout-important title="Instructions"}
You can count work on this week's TidyTuesday toward the exceptional work required for an A in the Homework component.
:::

Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}). Provide sufficient background for readers to grasp your narrative.

------------------------------------------------------------------------

\_\_\_\_

```{r}
library(tidyverse)
```

```{r}
# Using R
# Option 1: tidytuesdayR R package 
## install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2026-03-24')
## OR
tuesdata <- tidytuesdayR::tt_load(2026, week = 12)

pi_digits <- tuesdata$pi_digits

# Option 2: Read directly from GitHub

pi_digits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-03-24/pi_digits.csv')
```

```{r}
pi_statistics <- pi_digits %>%
  group_by(digit) %>%
  summarise(count = n()) %>%
  mutate(
    proportion = (count / sum(count)) * 100,
    deviation = count - 100000,
    Maximun = count == max(count),
    digit = as.factor(digit))
```

```{r}
summary(pi_statistics)
```

```{r}
pi_statistics |> 
  mutate(diff = proportion - 10) |> 
  ggplot(aes(x = factor(digit), y = diff)) +
  geom_col(fill = "skyblue", color = "black") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  scale_y_continuous(labels = function(x) paste0(x + 10, "%")) +
  labs(
    title = "Which number appears more times in PI?",
    subtitle = "Deviation from the uniform 10% distribution",
    x = "Digits from 0 to 9",
    y = "Proportion (%)",
    caption = "Data: TidyTuesday 2026-03-24 | Author: Patricia Escobar") +
  theme_minimal() +
  geom_text(aes(label = sprintf("%.2f%%", proportion), vjust = ifelse(diff >= 0, -0.5, 1.5)), size = 3.5)
```

I looked at which numbers appear more times in PI relative to an expected 10% uniform distribution. I discovered that the number that appears more times is 5, making up 10.04% of all the numbers, and the number that appears fewer times is 6, making up 9.95% of all the numbers in Pi.
